---
title: "Weather Extremes in Singapore 2023"
author: "Team Violetred"
date: "2024-05-24"
format: html
---

## Introduction
This report analyzes the weather extremes in Singapore for the year 2023, highlighting significant weather conditions across different regions.

```{r setup, include=FALSE}
# Load necessary libraries
library(readxl)
library(tidyverse)
library(lubridate)
```

# Load the data

```{r}
weather_data <- read_excel("full_year_weather_data.xlsx")
```

# View the first few rows of the data

```{r}
head(weather_data)
```

# Data cleaning steps
# Convert date columns to Date type

```{r}
weather_data <- weather_data %>%
  mutate(Date = as_date(Date, format = "%Y-%m-%d"))
```

# Summary of data to check for any inconsistencies

```{r}
summary(weather_data)
```

# Create the improved visualization
# Use ggplot2 or any other suitable library

# Example data frame for demonstration (replace with actual data structure)
```{r}
weather_extremes <- data.frame(
  Region = c("Woodlands", "Ang Mo Kio", "Pulau Ubin", "Pasir Panjang", "Newton", "Kallang", "East Coast Parkway", "East Coast Parkway"),
  Event = c("Wettest month", "Hottest day", "Coolest month", "Strongest wind gust", "Coolest day", "Wettest day", "Hottest month", "Driest month"),
  Measurement = c(569.4, 37, 26, 85, 21.1, 225.5, 29.8, 39.8),
  Date = c("December", "May 13", "January", "May 16", "March 1", "Feb 28", "May", "October"),
  Value = c(569.4, 37, 26, 85, 21.1, 225.5, 29.8, 39.8)
)

ggplot(weather_extremes, aes(x=Region, y=Measurement, fill=Event)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=c("blue", "red", "green", "orange", "purple", "brown", "yellow", "pink")) +
  labs(title="Weather Extremes in Singapore 2023",
       subtitle="Temperature, Rainfall, and Wind Gusts",
       x="Region",
       y="Measurement",
       fill="Event") +
  theme_minimal() +
  theme(legend.position="bottom")
```