---
title: "Weather Extremes in Singapore 2023"
author: "Team Violetred"
date: "2024-05-24"
format: html
---

## Introduction
This report analyzes the weather extremes in Singapore for the year 2023, highlighting significant weather conditions across different regions. 

The code below requires the following packages:

```{r}
#| label: setup
#| include: false

# Load necessary libraries
library(tidyverse)
library(readxl)
library(ggplot2)
library(ggrepel)
library(ggimage) # To add the icons above the labels
library(sf)
library(tibble)
library(dplyr)
library(geojsonsf)
library(scales)
library(viridis) # For color-blind friendly palette
library(ggspatial) # For north arrow and scale bar
library(leaflet) # For interactive maps
library(stringr) # For string manipulation
library(rvest) # For web scraping
```

## Loading the datasets
```{r}
weather_data <- read_excel("full_year_weather_data.xlsx")
sg_coords <- read_csv("singapore_city_coordinates_improved.csv")
```

## Cleaning the datasets
```{r}
# Add the year to the date entries and convert to datetime format
weather_data <- weather_data %>%
  mutate(Date = as.Date(paste(Date, "2023"), format = "%d %b %Y"))

# Convert Daily Rainfall Total to numeric, forcing non-numeric values to NA
weather_data <- weather_data %>%
  mutate(`Daily Rainfall Total (mm)` = as.numeric(`Daily Rainfall Total (mm)`))

# Remove rows with NA values in the Daily Rainfall Total column
weather_data <- weather_data %>%
  filter(!is.na(`Daily Rainfall Total (mm)`))
```

## Combining the datasets
```{r}
sg_coords <- sg_coords |>
  select(place = `Place`, city = `City`, y = `latitude`, x = `longitude`)

# Join on the "city" column
join_city <- weather_data %>%
  left_join(sg_coords, by = c("Location" = "city"))

# Join on the "place" column
join_place <- weather_data %>%
  left_join(sg_coords, by = c("Location" = "place"))


# Combine the results
combined_weather_data <- bind_rows(
  join_city %>% filter(!is.na(x) & !is.na(y)),
  join_place %>% filter(!is.na(x) & !is.na(y))
)

combined_weather_data
```

```{r}
# View the cleaned data
head(combined_weather_data)
```

## Total Rainfall Analysis
```{r}
weather_total_rainfall <- combined_weather_data |>
  select(location = `Location`, date = `Date`, `Daily Rainfall Total (mm)`, x, y) |>
  mutate(date = format(date, "%Y-%m")) |>
  summarise(total_rainfall = sum(`Daily Rainfall Total (mm)`, na.rm = TRUE), .by = c(location, date, x, y))

wettest <- weather_total_rainfall |>
  slice_max(total_rainfall, n = 10, with_ties = FALSE)

driest <- weather_total_rainfall |>
  slice_min(total_rainfall, n = 10, with_ties = FALSE)

wettest
driest
```

## Temperature Analysis
```{r}
weather_temperature <- combined_weather_data |>
  select(location = `Location`, date = `Date`, temp = `Mean Temperature (°C)`, x, y) |>
  mutate(date = format(date, "%Y-%m"), temp = as.numeric(temp)) |>
  summarise(mean_temperature = mean(temp, na.rm = TRUE), .by = c(location, date, x, y))

hottest <- weather_temperature |>
  slice_max(mean_temperature, n = 10, with_ties = FALSE)

coolest <- weather_temperature |>
  slice_min(mean_temperature, n = 10, with_ties = FALSE)

hottest
coolest
```

## Wind Gust Analysis
```{r}
weather_wind_speed <- combined_weather_data |> 
  select(location = `Location`, date = `Date`, wind_speed = `Max Wind Speed (km/h)`, x, y) |>
  mutate(date = format(date, "%Y-%m"), wind_speed = as.numeric(wind_speed))

weather_wind_speed <- weather_wind_speed |> 
  slice_max(wind_speed, n = 10, with_ties = FALSE)

weather_wind_speed
```

```{r}
combined_extremes <- bind_rows(
  wettest %>% mutate(event = "Wettest", amount = total_rainfall),
  driest %>% mutate(event = "Driest", amount = total_rainfall),
  hottest %>% mutate(event = "Temperature (Hot)", amount = mean_temperature),
  coolest %>% mutate(event = "Temperature (Cool)", amount = mean_temperature),
  weather_wind_speed %>% mutate(event = "Wind Speed", amount = wind_speed)
)

combined_extremes
```

## Preparing for Map Plot

```{r}
# Custom color palette
event_colors <- c(
  "Wettest" = "#5ab4ac",
  "Driest" = "#d8b365",
  "Temperature (Cool)" = "#4575b4",
  "Temperature (Hot)" = "#d73027",
  "Wind Speed" = "#91bfdb"
)

wettest_date <-
  combined_extremes |>
  slice_max(total_rainfall, n = 1, with_ties = FALSE) |>
  mutate(event = "Wettest", amount = total_rainfall)

driest_date <-
  combined_extremes |>
  slice_min(total_rainfall, n = 1, with_ties = FALSE) |>
  mutate(event = "Driest", amount = total_rainfall)

hottest_date <-
  combined_extremes |>
  slice_max(mean_temperature, n = 1, with_ties = FALSE) |>
  mutate(event = "Temperature (Hot)", amount = mean_temperature)

coolest_date <-
  combined_extremes |>
  slice_min(mean_temperature, n = 1, with_ties = FALSE) |>
  mutate(event = "Temperature (Cool)", amount = mean_temperature)

strongest_wind_gust <-
  combined_extremes |>
  slice_max(wind_speed, n = 1, with_ties = FALSE) |>
  mutate(event = "Wind Speed", amount = wind_speed)

labels <- bind_rows(
  wettest_date, 
  driest_date, 
  hottest_date, 
  coolest_date, 
  strongest_wind_gust
) |>
  mutate(
    icon = c(
      "icons/rainfall_icon.png",
      "icons/dry_icon.jpg",
      "icons/sun_icon.png",
      "icons/cool_temperature_icon.png",
      "icons/wind_icon.png"
    )
  )

labels
```

## Plotting Map
```{r}
# Read the GeoJSON file
singapore_map <- geojson_sf("sg.geojson")

# Create the plot
ggplot() +
  # Base map
  geom_sf(data = singapore_map, fill = "ivory", color = "gray80", size = 0.5) +
  
  # Events
  geom_count(
    data = combined_extremes,
    aes(x = x, y = y, color = event),
    alpha = 0.8
  ) +
  scale_size_area() +
  
  geom_label_repel(
    data = labels,
    aes(
      x = x, y = y,
      label = paste0(location, "\n", event, " (" , amount, ")", "\n", date)
    ),
    size = 2.5, box.padding = 0.5, point.padding = 0.3,
    segment.color = "grey50", force = 10, max.overlaps = Inf,
    fill = "white", color = "black"
  ) +
  
  # Theme and labels
  theme_minimal() +
  labs(
    title = "Weather Extremes in Singapore (2023)",
    subtitle = "Highlighting the most significant weather events across different locations",
    x = "Longitude",
    y = "Latitude",
    caption = "Map Source: Singapore GeoJSON"
  ) +
  
  # Color and size scales
  scale_fill_manual(values = event_colors, name = "Event Type") +
  scale_fill_fermenter(palette = "Blues", name = "Population Count", breaks = seq(0, 500000, by = 50000), labels = scales::comma, direction = 1) +
  scale_color_manual(values = event_colors, name = "Event Type") +
  scale_size_continuous(range = c(3, 10), name = "Months") +
  
  # Adjust theme elements
  theme(
    plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
    plot.subtitle = element_text(size = 12, color = "gray30", margin = margin(b = 20)),
    legend.position = "right",
    legend.box = "vertical",
    legend.margin = margin(t = 10, r = 0, b = 10, l = 0),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    text = element_text(size = 10),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  
  # Guides
  guides(
    color = guide_legend(override.aes = list(size = 5)),
    size = guide_legend(override.aes = list(color = "gray30"))
  ) +
  
  # Add north arrow and scale bar
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
    style = north_arrow_minimal
  ) +
  annotation_scale(location = "bl", width_hint = 0.3, pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in")) +
  
  # Set coordinate limits
  coord_sf(xlim = c(103.6, 104.1), ylim = c(1.15, 1.5), expand = FALSE)
```
## Map of waterbodies 
```{r}
waterbodies <- geojson_sf("waterbody.geojson")

# Check the structure of the GeoJSON file
print(st_geometry_type(waterbodies))
print(st_bbox(waterbodies))

# Ensure CRS is set correctly
st_crs(waterbodies) <- 4326

# Plot the waterbodies
ggplot() +
  geom_sf(data = waterbodies, fill = "blue") +
  ggtitle("Map of Waterbodies") +
  theme_minimal()
```
## Population Density by Planning Area
```{r}
# Load the population data from the Excel file
population_data_excel_path <- "respopagesexfa2023e.xlsx"
population_data <- read_excel(population_data_excel_path, skip = 2)

# Display the first few rows of the dataset
head(population_data)

# Rename columns
population_data <- population_data %>%
  rename(PA = `Planning Area`, Pop = `2023`)

# Convert Pop column to numeric, coercing any non-numeric values to NA
population_data <- population_data %>%
  mutate(Pop = as.numeric(Pop))

# Display the cleaned and renamed data
head(population_data)

# Select the first row of each distinct Planning Area, which contains the total population
population_aggregated <- population_data %>%
  group_by(PA) %>%
  slice(1) %>%
  ungroup()

# Display the aggregated population data
head(population_aggregated)
```

```{r}
# Merge population data with geographical data
singapore_map <- singapore_map %>%
  left_join(population_aggregated, by = c("planning_area" = "PA"))

# Plot the population density heatmap
ggplot() +
  geom_sf(data = singapore_map, aes(fill = Pop), color = "gray80", size = 0.2) +
  # Use viridis palette for perceptual uniformity and accessibility
  scale_fill_viridis_c(option = "plasma", name = "Population Count", labels = comma) +
  theme_minimal() +
  labs(
    title = "Population Density in Singapore (2023)",
    subtitle = "Aggregated by Planning Area",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: Singapore Department of Statistics"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
    plot.subtitle = element_text(size = 12, color = "gray30", margin = margin(b = 20)),
    legend.position = "right",
    legend.box = "vertical",
    legend.margin = margin(t = 10, r = 0, b = 10, l = 0),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    text = element_text(size = 10),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank()
  )
```


## Merging Weather Extremes Plot, Population Density Plot & Waterbodies Plot
```{r}
#| fig-width: 10
#| fig-height: 10

# Create the combined plot
ggplot() +
  # Base map for population density
  geom_sf(data = singapore_map, aes(fill = Pop), color = "gray80") +
  
  # Waterbodies
  geom_sf(data = waterbodies, fill = "blue", alpha = 0.5) +
  
  # Events data
  geom_count(
    data = combined_extremes,
    aes(x = x, y = y, color = event),
    alpha = 0.8
  ) +
  scale_size_area() +
  
  # Labels for weather events
  geom_label_repel(
    data = labels,
    aes(
      x = x, y = y,
      label = paste0(location, "\n", event, "\n", date)
    ),
    size = 2.5, box.padding = 0.5, point.padding = 0.3,
    segment.color = "grey50", force = 10, max.overlaps = Inf,
    fill = "white", color = "black"
  ) +
  
  # Color and size scales
  scale_color_manual(values = event_colors, name = "Event Type") +
  scale_size_continuous(range = c(3, 10), name = "Months") +
  scale_fill_fermenter(palette = "Blues", name = "Population Count", breaks = seq(0, 500000, by = 50000), labels = scales::comma, direction = 1) +
  
  # Theme and labels
  theme_minimal() +
  labs(
    title = "Weather Extremes in Singapore (2023)",
    subtitle = "Combining population density with significant weather events",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: Singapore Department of Statistics and Weather Data"
  ) +
  
  # Adjust theme elements
  theme(
    plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
    plot.subtitle = element_text(size = 12, color = "gray30", margin = margin(b = 20)),
    legend.position = "right",
    legend.box = "vertical",
    legend.margin = margin(t = 10, r = 0, b = 10, l = 0),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    text = element_text(size = 10),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  
  # Guides for color and size
  guides(
    color = guide_legend(override.aes = list(size = 5)),
    size = guide_legend(override.aes = list(color = "gray30"))
  ) +
  # Add north arrow and scale bar
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
    style = north_arrow_minimal
  ) +
  annotation_scale(location = "bl", width_hint = 0.3, pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in")) +
  # Set coordinate limits
  coord_sf(xlim = c(103.6, 104.1), ylim = c(1.15, 1.5), expand = FALSE)
```


## Adding Icons above the labels for the different weather extremes
```{r}
#| fig-width: 10
#| fig-height: 10

# Add icons to the map above
ggplot() +
  # Base map for population density
  geom_sf(data = singapore_map, aes(fill = Pop), color = "gray80", size = 0.2) +
  
  # Waterbodies
  geom_sf(data = waterbodies, fill = "blue", alpha = 0.5) +
  
  # Events data
  geom_count(
    data = combined_extremes,
    aes(x = x, y = y, color = event), 
    shape = 21 ,alpha = 0.8
  ) +
  
  # Icons for weather events
  geom_image(
    data = labels,
    aes(x = x, y = y, image = icon),
    size = 0.05, # Adjust size as needed
    by = "width" # Ensures the icon sizes are consistent
  ) +
  # Labels for weather events
  geom_label_repel(
    data = labels,
    aes(
      x = x, y = y,
      label = paste0(location, "\n", event, "\n", date)
    ),
    size = 2.5, box.padding = 0.5, point.padding = 0.3,
    segment.color = "grey50", force = 10, max.overlaps = Inf,
    fill = "white", color = "black", hjust = 0.5
  ) +
  # Color and size scales
  scale_fill_fermenter(palette = "Blues", name = "Population Count", breaks = seq(0, 500000, by = 50000), labels = scales::comma, direction = 1) +
  scale_color_manual(values = event_colors, name = "Event Type") +
  scale_size_continuous(range = c(10,20), name = "Months") +

  # Theme and labels
  theme_minimal() +
  labs(
    title = "Weather Extremes in Singapore (2023)",
    subtitle = "Adding icons to the combined map",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: Singapore Department of Statistics and Weather Data"
  ) +
  
  # Adjust theme elements
  theme(
    plot.title = element_text(size = 16, face = "bold", margin = margin(b = 10)),
    plot.subtitle = element_text(size = 12, color = "gray30", margin = margin(b = 20)),
    legend.position = "right",
    legend.box = "vertical",
    legend.margin = margin(t = 10, r = 0, b = 10, l = 0),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    text = element_text(size = 10),
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank()
  ) +
  
  # Guides for color and size
  guides(
    color = guide_legend(override.aes = list(size = 5)),
    size = guide_legend(override.aes = list(color = "gray30"))
  ) +
  
  # Add north arrow and scale bar
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
    style = north_arrow_minimal
  ) +
  
  annotation_scale(location = "bl", width_hint = 0.3, pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in")) +
  
  # Set coordinate limits
  coord_sf(xlim = c(103.6, 104.1), ylim = c(1.15, 1.5), expand = FALSE)
```
